---
title: "Chapter 7"
author: "BPR"
date: "8/23/2020"
output: word_document
---
Need to use qgraph to plot the adjacency matrices?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
{         # Put the cursor on this line, and run it; because
          #  of the curly braces, all the code within
          #  the braces will be run.
  c("data.table",   # fread() for FAST and flexible data input
    "dplyr",        # Data wrangling
    "dtplyr",       # For faster data wrangling with tidyverse structure
    "forecast",     # ARIMA modeling
#    "goftest",      # CVM test
    "here",         # To assist with folder structure
    "knitr",        # Better table printing
#    "MASS",         # For fitting distributions
# Something is wrong with nonlinearAnalysis & this
#  version of R. Hence, use the six tests individually.
#    "nonlinearAnalysis",
#    "poweRlaw") ->  # To fit power law data; MASS::fitdistr() can't
    "pracma",
    "qgraph",       # Plot transition networks.
    "tseriesChaos") ->
  package_names

  for (package_name in package_names) {
    if (!is.element(package_name, installed.packages()[, 1])) {
      install.packages(package_name)
    }
    library(package_name,
            character.only = TRUE,
            quietly = TRUE,
            verbose = FALSE
    )
  }
  rm(list = c("package_names", "package_name"))
}

# Reprodcibility
set_here()      # This helps with folder structure when shifting computers

# Usually, seeds for random number generation would be set here. They are, for
#  pedagogical reasons, set with each snippet of code.

# Stuff I prefer
options(show.signif.stars = FALSE) # Do not show significance stars! Showing
                                   #  significance stars encourages the
                                   #  conflation of significant and effect
                                   #  size.
options(digits = 5)                # Round to 5 digits
```

# Because something is going on the package::nonlinearAnalysis, the 6 tests of nonlinearityTests are available through these packages:
```{r setupExtra}
{         # Put the cursor on this line, and run it; because
          #  of the curly braces, all the code within
          #  the braces will be run.
  c("tseries",
    "TSA") ->
  package_names

  for (package_name in package_names) {
    if (!is.element(package_name, installed.packages()[, 1])) {
      install.packages(package_name)
    }
    library(package_name,
            character.only = TRUE,
            quietly = TRUE,
            verbose = FALSE
    )
  }
  rm(list = c("package_names", "package_name"))
}
```


```{r F7.1}
0:30 -> year
0.03 -> rate
1000 -> initial
initial * (1 + rate) ^ year -> population
plot(year, population,
     pch = 16,
     cex = 0.8,
     main = "Population Growth Example",
     xlab = "Time elapsed",
     ylab = "Population")
```


```{r F7.2a}
0:30 -> year
2.65 -> R

vector(mode = "numeric",
       length = 21) -> population
0.5 -> population[1]

for(generation in 1:(length(year)-1)) {
  R * population[generation] * (1 - population[generation]) ->
    population[generation + 1]
}
plot(year, population,
     pch = 16,
     cex = 0.8,
     main = "Population (with carrying capacity)",
     xlab = "Generation",
     ylab = "Fraction of /nCarrying Capacity",
     text(20, 0.6,
          paste("R =", R)))

```


```{r F7.2b}
0:30 -> year
3.513 -> R

vector(mode = "numeric",
       length = 21) -> population
0.5 -> population[1]

for(generation in 1:(length(year)-1)) {
  R * population[generation] * (1 - population[generation]) ->
    population[generation + 1]
}
plot(year, population,
     pch = 16,
     cex = 0.8,
     main = "Population (with carrying capacity)",
     xlab = "Generation",
     ylab = "Fraction of /nCarrying Capacity",
     text(20, 0.6,
          paste("R =", R)))

```


Figure 7.3(a) and 7.3(b)
 

```{r CS7.1}
0:30 -> year
0.03 -> rate
1000 -> initial
initial * (1 + rate) ^ year -> amount

auto.arima(ts(amount)) -> amount_arima
plot(forecast(amount_arima),       # Figure 7.3(a)
     main = "Compound Interest Forecast",
     xlab = "Year after deposit",
     ylab = "Amount in Account")

data("AirPassengers")
auto.arima(AirPassengers) -> air_arima
plot(forecast(air_arima),         # Figure 7.3(b)
     main = "Air Passengers",
     xlab = "Year",
     ylab = "Thousands of Passengers")

```

Nonlinearity test on cosine and the Lorenz.
```{r CS7.2}
# A cosine function
print("Terasvirta's neural network test")
print("Null Hypothesis: Mean is linear")
terasvirta.test(
  x=ts(cos(seq(0, 8*pi, pi/100))),  #  A cosine function
  type="Chisq")
print("White's neural network test")
print("Null Hypothesis: Mean is linear")
white.test(ts(cos(seq(0, 8*pi, pi/100))))
print("Keenan's test for nonlinearity")
print("Null Hypothesis: AR process")
Keenan.test(ts(cos(seq(0, 8*pi, pi/100))))$p.value
print("McLeod-Li test")
print("Null Hypothesis: ARIMA")
max(unlist(McLeod.Li.test(y = ts(cos(seq(0, 8*pi, pi/100))),
               plot = FALSE)))
print("Tsay test")
print("Null hypothesis: AR")
Tsay.test(ts(cos(seq(0, 8*pi, pi/100))))$p.value
print("Likelihood ratio test")
print("Null Hypothesis: AR")
print("Alternate Hypothesis: TAR process")
tlrt(ts(cos(seq(0, 8*pi, pi/100))))$p.value

# The magnitude of the Lorenz model
print("Terasvirta's neural network test")
print("Null Hypothesis: Mean is linear")
terasvirta.test(
  x=lorenz.ts,
  type="Chisq")
print("White's neural network test")
print("Null Hypothesis: Mean is linear")
white.test(lorenz.ts)
print("Keenan's test for nonlinearity")
print("Null Hypothesis: AR process")
Keenan.test(lorenz.ts)$p.value
print("McLeod-Li test")
print("Null Hypothesis: ARIMA")
max(unlist(McLeod.Li.test(y = lorenz.ts,
               plot = FALSE)))
print("Tsay test")
print("Null hypothesis: AR")
Tsay.test(lorenz.ts)$p.value
print("Likelihood ratio test")
print("Null Hypothesis: AR")
print("Alternate Hypothesis: TAR process")
tlrt(lorenz.ts)

```



```{r}
seq(from = 0,
    to = 8*pi,
    by = 0.01) -> theta
plot(theta, 1.5*cos(theta),
     type = 'l',
     xlab = "Time",
     ylab = "Position")
```


```{r ellipse}
# Fictitious data; this works for the figure
seq(from = 0,
    to = 2*pi,
    by = 0.01) -> theta
{
  plot(1.5* cos(theta), sin(theta),
       type = 'l',
       xlab = "Position",
       ylab = "Velocity")
  abline(v = 0)
  abline(h = 0)
  arrows(1.5/sqrt(2), 1/sqrt(2),
         1.5*cos(pi/4+0.01), sin(pi/4 + 0.01))
}


```


```{r F7.5}
# No header is given, so the default is "V1"; use that column.
fread("http://citadel.sjfc.edu/faculty/bricca/SCIPIE2019/logistic.txt",
  data.table = FALSE)$V1 -> logistic_series
plot(x = 1:(length(logistic_series)),
     y = logistic_series,
     xlab = "n",
     ylab = expression('x'[n]),
     xlim = c(500,700),  # Just look at part of it
     pch = 16,           # Use dots
     cex = 0.5)          # Use smaller dots
```


```{r F7.6}
seq(from = 0,
    to = 1,
    by = 0.1) -> n
seq(from = 0,
    to = 1,
    by = 0.1) -> np1
plot(n, np1,
     type = 'n',
     xlab = expression('x'[n]),   # [] puts a subscript in the label
     ylab = expression('x'[n+1]))


```


Figure 7.8
```{r CS7.3}
# No header is given, so the default is "V1"; use that column.
fread("http://citadel.sjfc.edu/faculty/bricca/SCIPIE2019/logistic.txt",
  data.table = FALSE)$V1 -> logistic_series
shift(logistic_series,
    1) -> xnp1   # "x n plus 1"
plot(xnp1, logistic_series,
     xlab = expression('x'[n]),   # [] puts in a subscript
     ylab = expression('x'[n+1]),
     pch = 16,
     cex = 0.5)


length(logistic_series) -> n
logistic_series[1:n-1] -> xn
logistic_series[2:n] -> xnp1
polyfit(xn, xnp1, 2) -> logistic_poly   # Fit a 2nd order polynomial

#{
#  plot(xn, xnp1,
#       col = "green",
#       xlab = expression('x'[n]),   # [] puts in a subscript
#       ylab = expression('x'[n+1]))
#  points(xn, 
#         logistic_poly[1]*xn*xn + logistic_poly[2]*xn + logistic_poly[3],
#     pch = 21,
#     cex = 0.4,
#     col = "blue")
#}

round(logistic_poly,5)

```

# Embed Lorenz
use of the mutual() and false.nearest() functions to determine the lag and the embedding dimension. Need a better example.
```{r CS7.4}
# Data?
# AMI and where there is a minimum. (Maybe use RJB or WBT data?)
mutual(lorenz.ts,       # works well. Do with and without plot.
       plot = FALSE) -> lorenz.ami
which.min(lorenz.ami)

false.nearest(lorenz.ts,
              m = 5,
              d = which.min(lorenz.ami),
              t = 100) -> fnn
which.min(fnn[1,]/fnn[2,])
```
Not surprisingly, this is a 1-dimensional embedding. (Try with the original Lorenz; do we get 3?)


Hmm, walking is also a 1-d thing...not surprising, as the delay isn't the important thing.

Figures 7.8 - 7.9
```{r CS7.5}
fread(file = "/Users/barneyricca/Dropbox/Dynamic analysis of gait and kicking/Data/Beginning Walk Full Term.csv",
      header = TRUE) -> walk1_df


# Figure 7.8(a) - 7.8(c)
plot(walk1_df$Time, walk1_df$ankleAngle,
     type = 'l')
plot(walk1_df$Time, walk1_df$kneeAngle,
     type = 'l')
plot(walk1_df$Time, walk1_df$hipFlex,
     type = 'l')

plot(walk1_df$ankleAngle[1:400], walk1_df$kneeAngle[1:400],
     type = 'l')


# OK, can do some DTW on these things.
# dtw(walk1_df$ankleAngle, walk3_df$ankleAngle) -> dtw1

```


```{r F7.12}

```



```{r F7.13}
fread("https://raw.githubusercontent.com/barneyricca/CSRforSS/master/Data/logistic.txt",
  data.table = FALSE)$V1 -> logistic_series
shift(logistic_series[1:30],
    1) -> xnp1   # "x n plus 1"
logistic_series[1:30] -> xn
plot(xnp1, xn,
     xlab = expression('x'[n]),   # [] puts in a subscript
     ylab = expression('x'[n+1]),
     type = 'l')
```


```{r F7.15}
walk <- function(secs) {
  runif(secs/2,
        min = 140,
        max = 180) -> an
  sin(pi*an/180) * sample(c(-1,1),
                          secs/2,
                          replace = TRUE) -> ve
  data.frame("Knee angle" = an,
             "Knee velocity" = ve) -> w_df
  return(w_df)
}

climb <- function(secs) {
  runif(secs/2,
        min = 130,
        max = 180) -> an
  1.2 * sin(pi*an/180) * sample(c(-1,1),
                                secs/2,
                                replace = TRUE) -> ve
  data.frame("Knee angle" = an,
             "Knee velocity" = ve) -> c_df
  return(c_df)
}

sit <- function(secs) {
  runif(secs/2,
        min = 89,
        max = 91) -> an
  runif(secs/2,
        min = -0.04,
        max = 0.04) -> ve
  data.frame("Knee angle" = an,
             "Knee velocity" = ve) -> si_df
  return(si_df)
}

stand <- function(secs) {
  runif(secs/2,
        min = 179,
        max = 180) -> an
  runif(secs/2,
        min = -0.04,
        max = 0.04) -> ve
  data.frame("Knee angle" = an,
             "Knee velocity" = ve) -> st_df
  return(st_df)
}

set.seed(715)
walk(1200) -> morn_df
rbind(morn_df, climb(120)) -> morn_df
rbind(morn_df, walk(1080)) -> morn_df
rbind(morn_df, sit(1200)) -> morn_df
rbind(morn_df, walk(10)) -> morn_df
rbind(morn_df, stand(240)) -> morn_df
rbind(morn_df, walk(10)) -> morn_df
rbind(morn_df, sit(3600)) -> morn_df
rbind(morn_df, walk(6)) -> morn_df
rbind(morn_df, stand(10)) -> morn_df
rbind(morn_df, walk(6)) -> morn_df
rbind(morn_df, sit(3600)) -> morn_df
rbind(morn_df, walk(6)) -> morn_df
rbind(morn_df, stand(10)) -> morn_df
rbind(morn_df, walk(6)) -> morn_df
rbind(morn_df, sit(3300)) -> morn_df

{
plot(morn_df$Knee.angle, morn_df$Knee.velocity,
     pch = 16,
     cex = 0.4)
lines(morn_df$Knee.angle, morn_df$Knee.velocity,
      lwd = 0.5)
}


{
plot(morn_df$Knee.angle, morn_df$Knee.velocity,
     pch = 16,
     cex = 0.4,
     xlim = c(88, 92),
     ylim = c(-0.05, 0.05))
lines(morn_df$Knee.angle, morn_df$Knee.velocity,
      lwd = 0.5)
}

{
plot(morn_df$Knee.angle, morn_df$Knee.velocity,
     pch = 16,
     cex = 0.4,
     xlim = c(175, 180),
     ylim = c(-0.05, 0.05))
lines(morn_df$Knee.angle, morn_df$Knee.velocity,
      lwd = 0.5)
}

```


```{r CS7.6}
matrix(c(0.9, 0.2, 0.1, 0.8),
       ncol = 2,
       nrow = 2,
       byrow = TRUE) -> pop_mat

eigen(pop_mat)$vectors[,1] -> pop_ss
# The first eigenvector is used because, well, that’s the one we want. If you want to know why, I refer you to a good linear algebra book.

# Normalize the vector so that the total percentage of people is equal to 1.
pop_ss / sum(pop_ss)

c(10000, 50000) ->initial_pops
pop_mat %*% initial_pops -> 
  pops                       # Initial populations
pop_mat %*% pops -> pops     # Year 2
pop_mat %*% pops -> pops     # Year 3
pop_mat %*% pops -> pops     # Year 4
pop_mat %*% pops -> pops     # Year 5
pop_mat %*% pops -> pops     # Year 6
pop_mat %*% pops -> pops     # Year 7
pop_mat %*% pops -> pops     # Year 8
pop_mat %*% pops -> pops     # Year 9
pop_mat %*% pops -> pops     # Year 10
pop_mat %*% pops -> pops     # Year 11
pop_mat %*% pops -> pops     # Year 12
pop_mat %*% pops -> pops     # Year 13
pop_mat %*% pops -> pops     # Year 14
pop_mat %*% pops -> pops     # Year 15
pop_mat %*% pops -> pops     # Year 16
pop_mat %*% pops -> pops     # Year 17
pop_mat %*% pops -> pops     # Year 18
pop_mat %*% pops -> pops     # Year 19
pop_mat %*% pops -> pops     # Year 20
pops


```



```{r CS7.7}
fread("https://raw.githubusercontent.com/barneyricca/CSRforSS/master/Data/ts_sample.txt",
  data.table = FALSE)$code -> data_series

unique(data_series) -> code_names           # Get the code names
length(data_series) -> n

# Now, make a transition matrix. The next command saves time:
matrix(0,
       nrow = length(code_names),
       ncol = length(code_names),
       dimnames = list(code_names, code_names)) -> # Name the rows & columns
  adj_mat

# Count the number of times one code is followed by another:
data_series[-1] -> shift_series
for(i in 1:(n-1)) {
  adj_mat[data_series[i], shift_series[i]] + 1 -> 
    adj_mat[data_series[i], shift_series[i]]
}
adj_mat                                   # Look at the result

# Make the matrix of transition probabilities (a.k.a., Markov matrix) from
#  the matrix of transition frequencies:
adj_mat/colSums(adj_mat) -> mark_mat
round(mark_mat,2)                                  # Look at the results

```
Equilibrium
```{r CS7.8}
eigenCodes <- eigen(t(mark_mat))
equilibrium <- eigenCodes$vectors[,1] / sum(eigenCodes$vectors[,1])
code_names -> names(equilibrium)
equilibrium
table(data_series) / length(data_series)
```

Bootstrap differences
```{r CS7.9}
fread("https://raw.githubusercontent.com/barneyricca/CSRforSS/master/Data/ts_sample.txt",
  data.table = FALSE)$code -> data_series

unique(data_series) -> code_names           # Get the code names

as.integer(length(data_series)/2) -> divider
data_series[1:divider] -> ds1
ds1[-1] -> shift_1

# Now, make a transition matrix. The next command saves time:
matrix(0,
       nrow = length(code_names),
       ncol = length(code_names),
       dimnames = list(code_names, code_names)) -> # Name the rows & columns
  adj_mat

# Count the number of times one code is followed by another:
for(i in 1:(divider-1)) {
  adj_mat[ds1[i], shift_1[i]] + 1 -> 
    adj_mat[ds1[i], shift_1[i]]
}
# Convert to probabilities, by row.
adj_mat/colSums(adj_mat) -> before_mat

data_series[(divider+1):length(data_series)] -> ds2
ds2[-1] -> shift_2
matrix(0,
       nrow = length(code_names),
       ncol = length(code_names),
       dimnames = list(code_names, code_names)) -> # Name the rows & columns
  adj_mat

#for(i in 1:(length(ds2)-1)) {
for(i in 1:1043) {
  adj_mat[ds2[i], shift_2[i]] + 1 -> 
    adj_mat[ds2[i], shift_2[i]]
}
adj_mat/colSums(adj_mat) -> after_mat

make_trans <- function(ts, code_names) {
  # Make a transition matrix fom data series ts. code_names contains
  #  all the possible codes, including those which do not appear in ts.
  #  (This makes the bootstrap comparisons easier when that comes up.)
  matrix(0,
         nrow = length(code_names),
         ncol = length(code_names),
         dimnames = list(code_names, code_names)) -> # Name the rows & columns
    adj_mat
  ts[-1] -> tss
  for(i in 1:(length(ts)-1)) {
    adj_mat[ts[i], tss[i]] + 1 ->
      adj_mat[ts[i], tss[i]]
  }
  return(adj_mat)
}

boot_trans <- function(mat1, mat2, probs, l1, R = 100, alpha = 0.05) {
  # mat1 and mat2 must have matching row & column names
  # mat1 and mat2 are Markov matrices
  # mat1 will be bootstrapped and compared to mat2
  # probs is the probability of each code appearing in the sub-sequence
  # l1 is the length of the first sub-sequence.
  # R is the number of replications
  
  vector(mode = "character", length = l1) -> boot_seq
  # Start the sub-sequence
  sample(colnames(mat1), 1, prob = probs) -> boot_seq[1]
  runif(1043) -> ps
  # Create the rest of the sub-seq
  # The next does NOT work!
  #  for(index in seq_along(2:l1)) {
  for(index in 2:l1) {
    sample(x = colnames(mat1),
           size = 1,
           prob = mat1[(boot_seq[index-1]),]) -> boot_seq[index]
  }
  
  boot_seq[-1] -> boot_shift
  
  # Create the transition matrix and compare the distance
  rep(0, R) -> boot_dist
  for(replicat in 1:R) {
    matrix(0, nrow = length(probs), 
           ncol = length(probs),
           dimnames = list(code_names, code_names)) -> boot_mat
    for(index in 1:(l1-1)) {
      boot_mat[boot_seq[index], boot_shift[index]] + 1 ->
        boot_mat[boot_seq[index], boot_shift[index]]
    }
    boot_mat/colSums(boot_mat) -> boot_mat
    sum((boot_mat-mat2)*(boot_mat-mat2)) -> boot_dist[replicat]
  }
  # Is there significance?
  if(sum((mat1-mat2)*(mat1-mat2)) < quantile(boot_dist, 1-alpha)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

boot_trans(before_mat, after_mat, 
           unname(table(ds1))/sum(table(ds1)), 
           divider)
```


```{r CS7.10}
# function trans_likely() for significance of transitions
```



embedd() function.


Markov:
```{r}
T = matrix(data=c(1/6, 1/3, 1/3, 1/6, 1/5, 1/5, 2/5, 1/5,
                  1/7, 1/7, 2/7, 3/7, 1/4, 1/8, 1/4, 3/8), nrow=4, ncol=4, byrow=TRUE)
N = 4000
chain = vector(mode="numeric", length=N)
chain[1] = sample(1:4, size=1)
for(i in 2:2000) {
  chain[i] = sample(1:4, 1, prob=T[chain[i-1],])
}

T2 = matrix(data=rep(0, times=16), nrow=4, ncol=4)
for(i in 1:(N-1)) {
  T2[chain[i], chain[i+1]] = T2[chain[i], chain[i+1]] + 1
}


T2 = T2 / rowSums(T2)
# Old version
#r_tot = rowSums(T2)
#for(i in 1:4) {
#  T2[i,] = T2[i,]/r_tot[i]
#}
T
T2
```



Appendix:
```{r}
# Modified from:
# https://desolve.r-forge.r-project.org/slides/tutorial.pdf
#install.packages("deSolve")
library(deSolve)

# Can we do mixed equations a la van Geert?
eqns <- function(t, y, parms) {
  -2 * y[2] * y[3] -> dy1
  -1.25 * dy1 * y[3] -> dy2
  -0.5 * y[1] * y[2] -> dy3
  return(list(c(dy1, dy2, dy3)))
}
# It appears so.

c(y1 = 1, y2 = 0.1, y3 = 0.9) -> yini
seq(0, 20, 0.01) -> times
ode(times = times,
    y = yini,
    func = eqns,
    parms = NULL) -> output_ode

head(output_ode, n = 3)

plot(output_ode)

#install.packages("scatterplot3d")
library(scatterplot3d)

par(mar = c(0,0,0,0))
scatterplot3d(output_ode[,-1],
              xlab = "",
              ylab = "",
              zlab = "",
              label.tick.marks = FALSE)

```


