---
title: "Chapter 4"
author: "BPR"
date: "6/16/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# The next five lines install the knitr package if necessary and add it to the 
#  environment. Just ignore these for now.
if (!is.element("knitr", 
                installed.packages()[, 1])) {  # If knitr isn't installed...
    install.packages("knitr")                  #  ...install it!
  }
library(knitr)                         # Add package:knitr to the environment

options(show.signif.stars = FALSE)     # Because significance stars are wrong! 
```

Figure 4.1. Testing the R installation.
```{r F4.1}
# This chunk contains the commands we used to test our R installation.
# You should copy and paste these lines, one at a time, into R.
c(1, 2, 3, 4, 5) -> integers
integers
mean(integers)
```

Figures 4.4 - 4.7: These come from a separate file that you create.

Code Snippet 4.1.
```{r CS4.1}
# The name in the previous line is for "Code Snippet 4.1." I'll follow this  
#  convention all the way through this document. "F4.n" will be for code used
#  to produce a figure, but which is not discussed in the text. "T4.n" will be
#  used for code that produces a table.

# Read some sample data, and store it in a variable. Notice that the name of 
#  the variable, "sample_data_df" describes the data (sample data) and 
#  describes the type of data ("data frame").
read.csv("https://raw.githubusercontent.com/barneyricca/CSRforSS/master/Data/sample.csv",                  # URL of data file
         header = TRUE) ->   # The first row has the column headers
  sample_data_df             # Name of the variable that stores the data
sample_data_df               # Display the data
lm(y ~ x,                    # Create a linear model (a.k.a. regression)
                             #  with y as the dependent variable and x
                             #  x as the independent variable.
   data = sample_data_df)    # Use x and y columns from sample_data_df
```

This code shows up as Figure 4.8. Sample code chunk.
```{r F4.8}
read.csv("https://raw.githubusercontent.com/barneyricca/CSRforSS/master/Data/sample.csv",                    # URL of data file
         header = TRUE) ->   # The first row has the column headers
  sample_data_df             # Name of the variable that stores the data
sample_data_df               # Display the data
lm(y~x,                      # Create a linear model (a.k.a. regression)
                             #  with y as the dependent variable and x
                             #  x as the independent variable.
   data = sample_data_df)    # Use x and y columns from sample_data_df
```

Code Snippet 4.2 and Figure 4.9
```{r CS4.2_F4.9}
1:10 -> x     # Put some data (the integers 1 to 10) into variable x.
x^2 -> y      # Put the square of x into y.
plot(x,y)     # Make a graph of y versus x.
```

Code Snippet 4.3 and Figure 4.10. A "better" plot.
```{r CS4.3_F4.10}
1:10 -> x                        # Put some data into variable x
x^2 -> y                         # Put the square of x into y
plot(x, y,
     main = "Dummy Data",        # Change the title above the plot
     sub = "More advanced plot", # Subtitle
     xlab = "X",                 # set the x-axis label
     ylab = "Y",                 # set the y-axis label
     cex.lab = 1.5,              # make the labels 50% bigger
     xlim = c(0,16),             # set the x-axis range 
     ylim = c(0,50),             # set the y-axis range
     pch = 16,                   # Use a solid dot, not an open circle
     cex = 0.8)                  # make the dots 80% of normal size
```

Code Snippet 4.4 and Figure 4.11
```{r CS4.4_F4.11}
rnorm(n = 1000,         # Draw 1000 times from a normal distribution
      mean = 0,         #  with a mean of zero
      sd = 1) ->        #  and a standard deviation of 1
  dummy                 #   and assign it to variable dummy
hist(dummy,             # Make a histogram of variable dummy
     breaks = 50,       #  using 50 bins
     main = "Normal distribution histogram")
```

Code Snippet 4.5 and Table 4.2.
```{r CS4.5_T4.2}
cbind(1:5, 6:10, 111:115) ->         # Combine 1:5, 6:10, and 111;111
                                     #  as three columns ("column bind")
  bad_table                          #  and store in the variable "bad_table".
bad_table                            # Print an ugly table

kable(x = bad_table,                 # Print bad_table
      format = "simple",             # No shading
      row.names = FALSE,             # No row names
      col.names = c("A", "B", "C"),  # Add (bold) column names
      align = 'c',                   # Center the numbers
      caption = "Output")            # Caption above the table
```

Code Snippet 4.6
```{r CS4.6}
2 + 3             # Addition (2 plut 3 = 5)
8 - 5             # Subtraction (8 minus 5 = 3)
8 * 7             # Multiplication (8 times 7 = 56)
16 / 2            # Division (16 divided by 2 = 8)
17 %/% 2          # Integer division quotient. 2 goes into 17 8 times
17 %% 2           # modulo, a.k.a. "remainder." 17/2 has remainder 1.
sqrt(81)          # Square root of 81
2 ^ 5             # Power (2 raised to the 5th power)
exp(3)            # Exponential
log(8, 2)         # log base 2 of 8
log(1000, 3)      # log base 3 of 1000
mean(1:10)        # Find the mean of the integers 1 to 10
sd(1:10)          # Find the standard deviation of the integers 1 to 10
median(1:10)      # Find the median of the integers 1 to 10
min(1:10)         # Find the minimum of the integers 1 to 10
max(1:10)         # Find the maximum of the integers 1 to 10
```

Code Snippet 4.7
```{r CS4.7}
1:15 -> data1          # Put the integers 1 through 15 into data1
seq(from = 2,          # Put a sequence, starting at 2...
    to = 30,           #  ...going up to 30...
    by = 2) ->         #  ...counting by twos...
  data2                #  ...into the variable data2
data1                  # Display data1 (i.e., 1 through 15)
data2                  # Display data2 (2, 4, 6, ... 30)
data1 + data2          # Add these together and display 
```

Code Snippet 4.8 and Figure 4.12(a)
```{r CS4.8_F4.12}
1:15 ->                        # Put the integers 1:15 into
  ind_var                      #  the independent variable
                   
3 * ind_var + 4 -> dep_var      # Use a linear relationship to create the
                                #  dependent variable
plot(ind_var, dep_var,          # Make a scatterplot
     main = "Noiseless data")   #  with a title.

set.seed(seed = 42)             # Look at the "Sampling" section in Chapter 4.
runif(n = 15,                   # Get 15 random numbers
      min = -5,                 #  from -5
      max = 5) +                #  to 5
  dep_var ->                    #  add them to the original
  dep_var                       #  and store the results back in
                                #  the original dependent variable.
plot(ind_var, dep_var,          # Make a scatterplot
     main = "Noisy data",       #  with a title and subtitle:
     sub = "Compare to Figure 4.12(a)")
```

Code Snippet 4.9, Tables 4.3 and 4.4, and Figure 4.13
```{r CS4.9_T4.3_T4.4_F4.13}
1:15 ->                             # Put the integers 1:15 into
  ind_var                           #  the independent variable
                   
3 * ind_var + 4 -> dep_var          # Use a linear relationship to create
                                    #  the dependent variable
set.seed(seed = 42)                 # Look at the "Sampling" section in 
                                    #  Chapter 4.
runif(n = 15,                       # Get 15 random numbers
      min = -5,                     #  from -5
      max = 5) +                    #  to 5
  dep_var ->                        #  add them to the original
  dep_var                           #  and store the results back in
                                    #  the original dependent variable.
plot(ind_var, dep_var,              # Scatterplot
     main = "Noiseless data")       #  with a title

lm(dep_var ~ ind_var) ->            # Regress dependent variable on the
                                    #  independent variable
  y_lm                              #  and store the results in y_lm
summary(y_lm)                       # Summarize the linear model (ugly)
kable(coefficients(summary(y_lm)),  # Nicer printing of the regression
                                    #  coefficients,
      digits = 3)                   #  rounded to 3 digits

{                                   # Put the scatterplot, line & text
                                    #  all on the same plot
  plot(ind_var, dep_var,            # Plot the data
       main = "Dummy Data",         # Change the title above the plot
       xlab = "Independent",        # set the x-axis label
       ylab = "Dependent",          # set the y-axis label
       xlim = c(0,16),              # set the x-axis range 
       ylim = c(0,50),              # set the y-axis range
       pch = 16,                    # Use a solid dot, not an open circle
       cex = 0.8)                   # make the dots 80% normal size
  abline(a = y_lm$coef[1],          # Add a line to the plot, with
                                    #  intercept from the linear model,
         b = y_lm$coef[2],          #  slope from the linear model,
         col = "darkgreen")         #  and using a dark green color.
  text(1, 40,                       # Put some text info on the plot @ (1,40)
       labels = paste(              # The label will have, pasted together,
         "slope =",                 #  "slope ="
         round(y_lm$coef[2],        #  and the value of the slope from the
                                    #  regression,
               digits = 2)),        #  rounded to 2 decimal places.
       adj = c(0,0),                # Lower left point at (1,40) on graph
       cex = 1.2)                   # 20% bigger text
  text(1, 35,                       # Put some more text on the plot
       labels = paste(              # Paste together
         "intercept = ",            #  "intercept = "
         round(y_lm$coef[1],        #  the value of the intercept from the
                                    #  regression
               digits = 2)),        #  rounded to two decimal places.
       adj = c(0,0),
       cex = 1.2)                   # 20% bigger text
}
```

Code Snippet 4.10 and Figure 4.14
```{r CS4.10_F4.14}
1:15 ->                             # Put the integers 1:15 into
  ind_var                           #  the independent variable
                   
3 * ind_var + 4 -> dep_var          # Use a linear relationship to create
                                    #  the dependent variable
set.seed(seed = 42)                 # Look at the "Sampling" section in 
                                    #  Chapter 4.
runif(n = 15,                       # Get 15 random numbers
      min = -5,                     #  from -5
      max = 5) +                    #  to 5
  dep_var ->                        #  add them to the original
  dep_var                           #  and store the results back in
                                    #  the original dependent variable.
plot(ind_var, dep_var,              # Scatterplot
     main = "Noiseless data")       #  with a title

lm(dep_var ~ ind_var) ->            # Regress dependent variable on the
                                    #  independent variable
  y_lm                              #  and store the results in y_lm

plot(y_lm,                          # Plot the results of the linear model.
                                    # Notice that the plot() command is
                                    #  pretty smart. It detects that y_lm
                                    #  is a linear model and plots
                                    #  appropriately!
     which = 1)                     # There are four plots; use the first.
                                    # Type ?plot.lm at the command prompt
                                    #  in the console, and press <Enter>
                                    #  to see what the others are.
```

Code Snippet 4.11
```{r CS 4.11}
read.csv(
  "https://raw.githubusercontent.com/barneyricca/CSRforSS/master/Data/ANOVA.csv",                                      # Read in some synthetic data;
      header = TRUE,                     # First row has the column names;
       stringsAsFactors = FALSE) ->      # Don't worry...just include this.
  anova_data                             # Store the data.

# Let's find the different in means between the groups, using the next
#  two commands. The second command takes a bit of unpacking, but it is
#  a useful trick. We use [] to get particular entries. In the first mean()
#  we only want the Score[] which have the corresponding Group equal to
#  "A". (Notice the double equal sign. That's important!) The second
#  mean() does the same thing, but uses Group "B".
print("Difference in means between groups A & B:")
mean(anova_data$Score[which(anova_data$Group == "A")]) -
  mean(anova_data$Score[which(anova_data$Group == "B")])

# ANOVA results:
summary(
  aov(Score ~ Group,                     # Standard ANOVA. The Score depends
                                         #  on the Group.
      data = anova_data))                # Print the ANOVA results

# Linear model results:
summary(
  lm(Score ~ Group,              # Do an ANOVA via a linear model!
     data = anova_data))         # Use the same data as above.

# Hey, look at that! The slope is equal to the difference in group means
#  and the p-value is the same as noted in the aov().
```

Code Snippet 4.12
```{r CS 4.12}
set.seed(412)
sample(LETTERS[1:4],                   # From the first four letters,
       size = 2,                       #  pick two,
       replace = FALSE,                #  do NOT replace between picks,
       prob = c(1/4, 1/4, 1/4, 1/4))   #  use equal probabilities.
## [1] "C" "A"

set.seed(412)
sample(LETTERS[1:4],                   # replace = FALSE and equal
       size = 2)                       #  probabilities are the defaults
## [1] "A" "D"                         # But the results are different!
#
#
# A note about the different results: Really, the default value in sample
#  is "prob = NULL" which is implemented in the underlying code slightly 
#  differently than listing the four equal probabilities. Look at this:
#
set.seed(412)
sample(LETTERS[1:4],                   # replace = FALSE and equal
       size = 2,                       #  probabilities are the defaults
       prob = NULL)                    # The default is "NULL."
## [1] "A" "D"                         # Now the results are the same.
```

Code Snippet 4.13 and Figures 4.15 and 4.16
```{r CS4.13_F4.15_F4.16}
set.seed(seed = 413)
runif(n = 1e5,          # Draw 100,000 samples uniformly
      min = -3,         # between -3...
      max = 3) ->       # ...and 3. Store the results...
  unif1                 # ...in variable unif1.
hist(unif1,             # Make a histogram of the data in variable unif1.
     breaks = 50,       # Divide the data into 50 bins
     xlim = c(-3, 15),  # Draw graph from x = -3 to x = 15. (See next example.)
     main = "Uniform Distribution [-3,3]")  # Put a title on the graph

runif(n = 1e5,
      min = 3,
      max = 15) -> # Use a different maximum value
  unif2
hist(unif2,
     breaks = 50,
     main = "Uniform Distribution [3,15]")

rnorm(n = 1e5,           # Draw 100,000 samples from a normal distribution...
      mean = 0,          # ...with a mean of 0...
      sd = 3) ->         # ...and standard deviation of 3. Store the results...
  norm1                  # ...in variable norm1.
hist(norm1,              # Make a histogram.
     breaks = 50,        #  
     main = expression(  # "expression" allows for math and Greek characters.
       paste(            # "paste" all this stuff together:
         "Normal Distribution: ",  # String
         mu,                       # Greek letter mu (because of expression)
         " = 0, ",                 # Another string
         sigma,                    # Greek letter sigma
         " = 1 ")),                # Another string
     xlim = c(-10, 20))            # Set the limits (see the next two)

rnorm(n = 1e5,
      mean = 4,
      sd = 3) -> 
  norm2
hist(norm2,
     breaks = 50,
     main = expression(paste("Normal Distribution: ",
                  mu,
                  " = 4, ",
                  sigma,
                  " = 3 ")),
     xlim = c(-10, 20))

rnorm(n = 1e5,
      mean = 4,
      sd = 1) -> 
  norm3
hist(norm3,
     breaks = 50,
     main = expression(paste("Normal Distribution: ",
                  mu,
                  " = 4, ",
                  sigma,
                  " = 1 ")),
     xlim = c(-10, 20))
```


Code Snippet 4.14: Fisher's experiment
```{r CS4.14}
set.seed(414)
8 -> cups            # Fisher proposed 8 cups
cups/2 -> cd2        # Half the cups will be tea and half milk
# First, randomly set up the cups, "M" for milk-first, "T" for
#  tea-first.
sample(                # Create a random sample...
  rep(c("M","T"),      #  ...from a collection of Milk and Tea...
      each = cd2),     #  ...using 4 of each.
  cups,                # Get a total of 8 cups.
  replace = FALSE) ->  # Don't reuse the cups. 
  treatment            # Put that into the variable treatment

1e6 -> N               # Repeat the experiment 1 million times

# First, it will make life easier to create a "function." The next command
#  allows us to treat "guess1" just like any other built-in R function.
#  (In other words, we are expanding R's capabilities.)
guess1 <- function(treat) {          # Define a function.
  # The function "guess1" takes the arrangement of cups (the variable treat),
  #  makes a random guess and returns how many cups were correctly identified
  #  by the guess.

  length(treat) -> cups   # How many cups are in the treatment?
  cups / 2 -> cd2         # Half the cups are milk, half are tea.

  sample(                 # This sample caommand should look familiar!
    rep(c("M","T"),       # Hint: Look up about a dozen lines
        each = cd2),
    cups,                          
    replace = FALSE) -> 
    attempt               # The guess is stored in the variable attemp

  # The next takes a bit of unpacking. The which() command returns the
  #  positions of all those entries that meet some criteria. In this case,
  #  the criteria is "treat == attempt"; NOTICE THE DOUBLE EQUAL SIGN!
  # The == means "if the first thing is equal to the second" then the
  #  criteria is met. (In computer science speak, we'll say "return TRUE.")
  # length() counts how many things met which()'s criteria.
  # Whew!
  return(                 # Return the results to whatever called the function
    length(
      which(treat == attempt)))
}

# The next is part 1, the large number of guesses. (See the text.)
replicate(N,                      # Do the next line N times
          guess1(treatment)) ->   # Make the guess!
  results                         # Store in a variable called results.

table(results) -> results_tab     # Create a table of the results
                                  # These will be in order from 0 correct
                                  #  up to all correct.

# Now, for the answer: What fraction of results guessed all the cups
#  correctly? 
results_tab[length(results_tab)] / N
```

Figure 4.16. Indicator of an error. This next chunk intentionally has an error in it, so you need to  correct the error (or remove the line) before knitting.
```{r F4.16}
# To knit this file, you would need to use the correct version of the next
#  command. Otherwise, the error will cause the knitting to stop.
# Hence, before knitting, put a hashtag in front of the next command.

lm(y ~ x, data sample_data_df)

```
